\bfseries Convolution-based nonlinear functions on tensor fields and tensor images

### 2022-08-29 paper/main.tex

In practice, a complete set of invariant $k$-$p$-tensor filters (filters $C$ such that $g\cdot C(g^{-1} \cdot a') = C(a')$) can be found by group averaging.
This works by building any set of linearly independent $k'$-$p'$-tensor filters (randomly or systematically), and then building, for each of these linearly independent filters $C$, a group-averaged filter $\bar{C}$ by
\begin{equation}
    \bar{C}(a') = \frac{1}{|B_d|}\,\sum_{g\in B_d} g\cdot C(g^{-1}\cdot a') ~,
\end{equation}
where $|B_d|$ is the number of group elements.
The averaged filters are then reduced to an ``eigen-set'' of orthogonal filters by the singular value decomposition.
In detail we started this averaging with a set of independent $k'$-$p'$-tensor convolution filters that span the space of all possible geometric convolution filters at side length $M$ ($M=2\,m+1$) and dimension $d$.
There are $M^d\,d^{k'}$ independent filters in this set.
The results of these operations are shown for $d=2$ and $m\in[1, 2]$ in \figref{fig:filters23} and \figref{fig:filters25}.
HOGG: Give a table saying how many filters there are for each dimension $d$ and each side length $M$ at each order $k$ and each parity $p$.
\begin{figure}[tp]
  \begin{mdframed}
  \color{captiongray}
  \begin{center}
\includegraphics[width=\textwidth]{notebooks/filter+0_2_3.png}\\
\includegraphics[width=\textwidth]{notebooks/filter+1_2_3.png}\\
\includegraphics[width=\textwidth]{notebooks/filter-1_2_3.png}
\includegraphics[width=\textwidth]{notebooks/filter+2_2_3.png}\\
\includegraphics[width=\textwidth]{notebooks/filter-2_2_3.png}
  \end{center}
\caption{All the filters for $d=2$, $m=1$ ($M=3$).
Notes: Scalars and pseudo-scalars are shown with signed colors; where there is no symbol in the box the value is zero. The $2$-$p'$-tensor filters are shown via the action of the tensor on an image of a letter ``R''; the transformation properties of the $2$-$p'$-filters are such that these filters may not look obviously invariant to rotations but they are.
There are no invariant pseudoscalar filters available at $(d,m) = (2,1)$.
Three of the $2$-$0$-tensors are duplicates of the $0$-$0$-tensor filters but multiplied by the identity tensor.\label{fig:filters23}}
  \end{mdframed}
\end{figure}
\begin{figure}[tp]
  \begin{mdframed}
  \color{captiongray}
  \begin{center}
\includegraphics[width=\textwidth]{notebooks/filter+0_2_5.png}\\
\includegraphics[width=\textwidth]{notebooks/filter-0_2_5.png}\\
\includegraphics[width=\textwidth]{notebooks/filter+1_2_5.png}\\
\includegraphics[width=\textwidth]{notebooks/filter-1_2_5.png}
  \end{center}
\caption{All the filters for $d=2$, $m=2$ ($M=5$).
Notes: Symbols and colors as in \figref{fig:filters23}.
We don't show the $2$-$p'$-tensor filters at $m=2$ because there are many of them.\label{fig:filters25}}
  \end{mdframed}
\end{figure}

>>>>

In practice, a complete set of invariant $k'$-$p'$-tensor filters (filters $C$ such that $g\cdot C(g^{-1} \cdot a') = C(a')$) can be found by group averaging.
This works by building any set of linearly independent $k'$-$p'$-tensor filters (randomly or systematically), and then building, for each of these linearly independent filters $C$, a group-averaged filter $\bar{C}$ by
\begin{equation}
    \bar{C}(a') = \frac{1}{|B_d|}\,\sum_{g\in B_d} g\cdot C(g^{-1}\cdot a') ~,
\end{equation}
where $|B_d|$ is the number of group elements (8 for $d=2$ and 27 for $d=3$).
In detail we took a systematic (not random) approach that initialized the averaging with a set of independent $k'$-$p'$-tensor convolution filters that span the space of all possible geometric convolution filters at side length $M$ ($M=2\,m+1$) and dimension $d$.
There are $M^d\,d^{k'}$ independent filters in this initial set.
The averaged filters are then reduced to an ``eigen-set'' of orthogonal, non-zero filters by the singular-value decomposition.
The singular-value decomposition reduces the number of filters dramatically.
The outputs of the singular-value decomposition can be normalized however seems appropriate.
In detail we normalized to get unit tensor norms in the non-zero pixels.
The results of these operations are shown for $d=2$ and $m\in[1, 2]$ in \figref{fig:filters23} and \figref{fig:filters25}.
HOGG: Give a table saying how many filters there are for each tuple of $(d,m,k',p')$.
HOGG: Perhaps remind the reader that the code is available.
\begin{figure}[tp]
  \begin{mdframed}
  \color{captiongray}
  \begin{center}
\includegraphics[width=\textwidth]{notebooks/filter_2_3.png}
  \end{center}
\caption{All the filters for $d=2$, $m=1$ ($M=3$).
Notes: Scalars and pseudo-scalars are shown with signed colors; where there is no symbol in the box the value is zero. The $2$-$p'$-tensor filters are shown via the action of the tensor on an image of a letter ``R''; the transformation properties of the $2$-$p'$-filters are such that these filters may not look obviously invariant to rotations, but they are.
There are no invariant pseudoscalar ($0$-$1$-tensor) filters available at $(d,m) = (2,1)$.
Three of the $2$-$0$-tensors are duplicates of the $0$-$0$-tensor (scalar) filters but multiplied by the identity tensor; there is only one non-trivial invariant $2$-$0$-tensor filter.
We don't show the $k'$-$p'$-filters at $k'>2$ because we don't know how to visualize them (even the $k'=2$ case is sketchy).\label{fig:filters23}}
  \end{mdframed}
\end{figure}
\begin{figure}[tp]
  \begin{mdframed}
  \color{captiongray}
  \begin{center}
\includegraphics[width=\textwidth]{notebooks/filter_2_5.png}
  \end{center}
\caption{All the filters for $d=2$, $m=2$ ($M=5$).
Notes: Symbols and colors as in \figref{fig:filters23}.
We don't show the $2$-$p'$-tensor filters at $m=2$ because there are many of them.\label{fig:filters25}}
  \end{mdframed}
\end{figure}
